---
title: "eda-q1"
output: pdf_document
date: "2025-04-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
ev_data <- read.csv("../Dataset/pre-processed-EV-dataset.csv")
```

```{r}
# Aggregate data by County
ev_county <- ev_data %>%
  group_by(county) %>%
  summarise(EV_Count = n())

# Aggregate data by City
ev_city <- ev_data %>%
  group_by(city) %>%
  summarise(EV_Count = n())

# Aggregate data by Legislative District
ev_district <- ev_data %>%
  group_by(legislative_district) %>%
  summarise(EV_Count = n())
```

```{r}
ev_district
```

```{r}
library(dplyr)
library(ggplot2)

top_counties <- ev_data %>%
  count(county, name = "EV_Count") %>%
  slice_max(EV_Count, n = 20)  # Select top 20 counties with highest EV count

ggplot(top_counties, aes(x = reorder(county, EV_Count), y = EV_Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip for better readability
  labs(title = "Top 20 Counties with Highest EV Adoption", x = "County", y = "EV Count") +
  theme_minimal()
```


```{r}
library(dplyr)
library(ggplot2)

top_cities <- ev_data %>%
  count(city, name = "EV_Count") %>%
  slice_max(EV_Count, n = 20)  # Select top 20 cities with highest EV count

ggplot(top_cities, aes(x = reorder(city, EV_Count), y = EV_Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip for better readability
  labs(title = "Top 20 Cities with Highest EV Adoption", x = "City", y = "EV Count") +
  theme_minimal()
```

```{r}
library(dplyr)
library(ggplot2)

top_ld <- ev_data %>%
  count(legislative_district, name = "EV_Count") %>%
  slice_max(EV_Count, n = 20)  # Select top 20 legislative_district with highest EV count

ggplot(top_ld, aes(x = reorder(legislative_district, EV_Count), y = EV_Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip for better readability
  labs(title = "Top 20 Cities with Highest EV Adoption", x = "City", y = "EV Count") +
  theme_minimal()
```




```{r}
########   Check data variables available in 'tidycensus'     ###########

library(tidycensus)

# Load list of variables for 2020 ACS 5-year
acs_vars <- load_variables(2020, "acs5", cache = TRUE)

# View sample
# View(acs_vars)  # Opens a searchable table in RStudio
```


##################             2                 ###################
```{r}
library(tidycensus)
library(tidyverse)

# Define variable codes (no friendly names)
vars <- c(
  "B01003_001",  # Total Population
  "B23025_003",  # Employed Population
  "B15003_022",  # Bachelor's Degree or Higher
  "B19013_001"   # Median Household Income
)

# Fetch data (geometry included)
census_data <- get_acs(
  geography = "tract",
  variables = vars,
  state = "WA",
  year = 2020,
  geometry = TRUE,
  output = "wide"  # Automatically pivots to wide format
)

census_data_clean <- census_data %>%
  rename(
    population = B01003_001E,
    employed_population = B23025_003E,
    bachelors_or_higher = B15003_022E,
    median_income = B19013_001E
  ) %>%
  select(GEOID, NAME, population, employed_population, 
         bachelors_or_higher, median_income, geometry)

glimpse(census_data_clean)
```



API key - c00131b3556a74bad78bbbb2d648afc4603bd046

```{r}
library(sf)
library(tidycensus)
library(stringr)

# Get Washington State county geometries
census_api_key("c00131b3556a74bad78bbbb2d648afc4603bd046")  # Set your Census API key
county_shapes <- get_acs(
  geography = "county",
  variables = "B01001_001",  # Total population (for reference)
  state = "WA",
  geometry = TRUE
)

# Step 1: Clean the county names in 'county_shapes' by removing 'County, Washington'
county_shapes <- county_shapes %>%
  mutate(NAME_clean = str_remove(NAME, " County, Washington"))

# Merge EV adoption data with county geometries
county_shapes <- county_shapes %>%
  left_join(ev_county, by = c("NAME_clean" = "county"))

# Step 1: Calculate centroids for each county (for positioning the text labels)
county_shapes <- county_shapes %>%
  mutate(centroid = st_centroid(geometry))

# Step 2: Extract the x and y coordinates of the centroids for text placement
county_shapes <- county_shapes %>%
  mutate(centroid_x = st_coordinates(centroid)[, 1],
         centroid_y = st_coordinates(centroid)[, 2])

# Step 3: Plot choropleth map with labels
ggplot(county_shapes) +
  geom_sf(aes(fill = EV_Count)) +
  scale_fill_gradient(low = "lightblue", high = "blue", na.value = "grey") +
  labs(title = "EV Adoption by County in Washington State", fill = "EV Count") +
  theme_minimal() +
  
  # Add EV_Count labels at the centroids of each county
  geom_text(aes(x = centroid_x, y = centroid_y, label = EV_Count), 
            size = 3,  # Adjust label size
            color = "black",  # Adjust text color
            check_overlap = TRUE)  # Prevent overlap of labels

```




```{r}
ev_data_modified_census_tract <- ev_data %>%
  mutate(census_tract = as.character(census_tract))
```

```{r}
# Prepare EV data - count vehicles by census tract
ev_counts <- ev_data_modified_census_tract %>%
  # Ensure census tract is character type
  mutate(census_tract = as.character(census_tract)) %>%
  # Count EVs per census tract
  group_by(census_tract) %>%
  summarize(EV_Count = n())
```

```{r}
# Merge the datasets
merged_data <- ev_counts %>%
  left_join(census_data_clean, by = c("census_tract" = "GEOID"))

# Verify the merged data structure
# glimpse(merged_data)

# Correlation analysis with the correct columns
cor_results <- cor(
  merged_data %>% select(EV_Count, population, employed_population, bachelors_or_higher, median_income),
  use = "pairwise.complete.obs"
)

print(cor_results)
```

#############  
Top 10 and bottom 10 census tracts' EV count

```{r}
library(tidyverse)
library(tidycensus)
library(sf)
library(leaflet)

summary(merged_data$EV_Count)

ev_census_overlay <- merged_data %>%
  mutate(
    EV_Count = replace_na(EV_Count, 0),  # Replace NA with 0 for visualization
    EV_Count = as.numeric(EV_Count)      # Ensure numeric type
)

# class(ev_census_overlay)


# Create color palette based on actual data range
pal <- colorNumeric(
  palette = "viridis",
  domain = ev_census_overlay$EV_Count
)

top_10 <- ev_census_overlay %>%
  arrange(desc(EV_Count)) %>%
  slice_head(n = 10)

bottom_10 <- ev_census_overlay %>%
  filter(EV_Count > 0 | is.na(EV_Count) == FALSE) %>%  # Optional: remove NA
  arrange(EV_Count) %>% # Keep only one row per value
  slice_head(n = 10)

# Transform CRS
ev_census_overlay <- st_as_sf(ev_census_overlay)
ev_census_overlay <- st_transform(ev_census_overlay, crs = 4326)
ev_census_overlay <- ev_census_overlay %>%
  mutate(
    fill_color = case_when(
      census_tract %in% top_10$census_tract ~ "green",
      census_tract %in% bottom_10$census_tract ~ "red",
      !is.na(EV_Count) ~ pal(EV_Count),
      TRUE ~ "#cccccc"  # fallback grey for any edge case
    )
  )

# Generate the map
leaflet(ev_census_overlay %>% st_transform(4326)) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~fill_color,
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",  # Display the tract name
      "<strong>EV Count:</strong>", ifelse(is.na(EV_Count), "No Data", EV_Count)
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~EV_Count,
    title = "EV Count per Tract",
    position = "bottomright"
  ) %>%
  addLegend(
    colors = c("green", "red"),
    labels = c("Top 10 EV Tracts", "Bottom 10 EV Tracts"),
    title = "Highlighted Tracts",
    position = "bottomleft"
  )
```

```{r}
pop_pal <- colorNumeric(
  palette = "Blues", 
  domain = ev_census_overlay$population
)

income_pal <- colorNumeric(
  palette = "YlOrRd", 
  domain = ev_census_overlay$median_income
)

# 6. Create the map with both Population and Income Data
leaflet(ev_census_overlay %>% st_transform(4326)) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~income_pal(median_income),  # Color by median income
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",  # Display the tract name
      "<strong>EV Count:</strong>", EV_Count, "<br>",
      "<strong>Population:</strong>", population, "<br>",
      "<strong>Median Income:</strong>", median_income
    )
  ) %>%
  addLegend(
    pal = income_pal,
    values = ~median_income,
    title = "Median Income per Tract",
    position = "bottomleft"
  )
```


```{r}
library(RColorBrewer)
ev_pal <- colorNumeric(
  palette = "Blues", 
  domain = ev_census_overlay$EV_Count
)


pop_pal <- colorNumeric(
  palette = brewer.pal(9, "YlGn"), 
  domain = ev_census_overlay$population
)

income_pal <- colorNumeric(
  palette = "YlOrRd", 
  domain = ev_census_overlay$median_income
)

emp_pop_pal <- colorNumeric(
  palette = brewer.pal(9, "YlGn"), 
  domain = ev_census_overlay$employed_population
)

bach_pal <- colorNumeric(
  palette = "OrRd", 
  domain = ev_census_overlay$bachelors_or_higher
)

# Generate the map with layer controls for Population, Income, and EV Count
leaflet(ev_census_overlay) %>%
  addTiles() %>%

  # Add Population Layer
  addPolygons(
    fillColor = ~pop_pal(population),
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>Population:</strong>", population
    ),
    group = "Population"
  ) %>%

  # Add Median Income Layer
  addPolygons(
    fillColor = ~income_pal(median_income),
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>Median Income:</strong>", median_income
    ),
    group = "Median Income"
  ) %>%
  
  
  # Add Employed Population Layer
  addPolygons(
    fillColor = ~emp_pop_pal(employed_population),
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>Employed Population:</strong>", employed_population
    ),
    group = "Employed Population"
  ) %>%

  # Add Bachelors or Higher Layer
  addPolygons(
    fillColor = ~bach_pal(bachelors_or_higher),
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>Bachelors or Higher:</strong>", bachelors_or_higher
    ),
    group = "Bachelors or Higher"
  ) %>%
  

  # Add EV Count Layer
  addPolygons(
    fillColor = ~ev_pal(EV_Count),
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>EV Count:</strong>", EV_Count
    ),
    group = "EV Count"
  ) %>%

  # Add Legends for Each Layer
  addLegend(
    pal = pop_pal,
    values = ~population,
    title = "Population",
    position = "bottomleft",
    group = "Population"
  ) %>%
  
  addLegend(
    pal = income_pal,
    values = ~median_income,
    title = "Median Income",
    position = "bottomright",
    group = "Median Income"
  ) %>%
  
  # addLegend(
  #   pal = emp_pop_pal,
  #   values = ~employed_population,
  #   title = "Employed Population",
  #   position = "bottomleft",
  #   group = "Employed Population"
  # ) %>%
  
  addLegend(
    pal = bach_pal,
    values = ~bachelors_or_higher,
    title = "Bachelor or Higher",
    position = "topleft",
    group = "Bachelor or Higher"
  ) %>%

  addLegend(
    pal = ev_pal,
    values = ~EV_Count,
    title = "EV Count per Tract",
    position = "topright",
    group = "EV Count"
  ) %>%

  # Add Layer Controls to Toggle Between Layers
  addLayersControl(
    baseGroups = c("Population", "Median Income", "Bachelor or Higher", "Employed Population", "EV Count"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


```{r}
intersection_layer <- ev_census_overlay %>%
  filter(population > 0, median_income > 0, EV_Count > 0, employed_population > 0, bachelors_or_higher > 0)

# Create a combined score for visualization
intersection_layer <- intersection_layer %>%
  mutate(
    combined_score = (population / max(population, na.rm = TRUE)) +
                     (median_income / max(median_income, na.rm = TRUE)) +
                     (EV_Count / max(EV_Count, na.rm = TRUE)) +
                      (employed_population / max(employed_population, na.rm = TRUE)) +
                      (bachelors_or_higher / max(bachelors_or_higher, na.rm = TRUE))
  )

# Identify top 10 and bottom 10 tracts based on combined score
top_10_tracts <- intersection_layer %>%
  arrange(desc(combined_score)) %>%
  slice_head(n = 10)

bottom_10_tracts <- intersection_layer %>%
  filter(combined_score > 0 | is.na(combined_score) == FALSE) %>% # Ensure we don't include zeros
  arrange(combined_score) %>%
  slice_head(n = 10)


# 7. Add highlight category
intersection_layer <- intersection_layer %>%
  mutate(highlight_category = case_when(
    census_tract %in% top_10_tracts$census_tract ~ "Top 10",
    census_tract %in% bottom_10_tracts$census_tract ~ "Bottom 10",
    TRUE ~ "Other"
  ))

# Define color palettes
# Top 10 will be green, Bottom 10 will be red, Others will use the PuBu gradient
highlight_pal <- colorFactor(
  palette = c("red", "green"),  # Make "Other" gray to show underlying color
  domain = c("Top 10", "Bottom 10")
)

# Create color palette for the intersection layer (for all tracts)
intersection_pal <- colorNumeric(
  palette = "PuBu", 
  domain = intersection_layer$combined_score
)

# 9. Generate the map
leaflet(intersection_layer) %>%
  addTiles() %>%
  
  addPolygons(
    fillColor = ~intersection_pal(combined_score),
    color = "#444444",
    weight = 1,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>Population:</strong>", population, "<br>",
      "<strong>Median Income:</strong>", median_income, "<br>",
      "<strong>EV Count:</strong>", EV_Count, "<br>",
      "<strong>Combined Score:</strong>", round(combined_score, 3)
    ),
    group = "All Tracts"
  ) %>%
  
  addPolygons(
    data = intersection_layer %>% filter(highlight_category != "Other"),
    fillColor = ~highlight_pal(highlight_category),
    color = "black",
    weight = 2,
    smoothFactor = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    popup = ~paste(
      "<strong>Census Tract:</strong>", NAME, "<br>",
      "<strong>Population:</strong>", population, "<br>",
      "<strong>Median Income:</strong>", median_income, "<br>",
      "<strong>Employed Population:</strong>", employed_population, "<br>",
      "<strong>Bachelor or Higher:</strong>", bachelors_or_higher, "<br>",
      "<strong>EV Count:</strong>", EV_Count, "<br>",
      "<strong>Combined Score:</strong>", round(combined_score, 3),
      "<strong>Category:</strong>", highlight_category
    ),
    group = "Highlighted Tracts"
  ) %>%
  
  addLegend(
    pal = intersection_pal,
    values = ~combined_score,
    title = "Combined Score",
    position = "bottomright"
  ) %>%
  
  addLegend(
    colors = c("green", "red"),
    labels = c("Top 10 Tracts", "Bottom 10 Tracts"),
    title = "Highlighted Tracts",
    position = "bottomleft"
  ) %>%
  
  addLayersControl(
    overlayGroups = c("All Tracts", "Highlighted Tracts"),
    options = layersControlOptions(collapsed = FALSE)
  )

```

<<<<<<< Updated upstream





###.     Optional, if it does not provide satisfactory results then we'll remove it
=======
>>>>>>> Stashed changes
###      Spatial Regression    ###

```{r}
library(spdep)
library(spatialreg)
library(sf)

# Ensure your data is properly formatted
analysis_data <- intersection_layer %>% 
  st_drop_geometry() %>% 
  select(EV_Count, population, median_income, employed_population, bachelors_or_higher) %>%
  na.omit()

# Convert to spatial object
spatial_data <- st_as_sf(intersection_layer) %>% 
  filter(!is.na(EV_Count), !is.na(population), !is.na(median_income))
```


```{r}
# Create neighbors list
coords <- st_centroid(st_geometry(spatial_data))
knn_neighbors <- knn2nb(knearneigh(coords, k = 5))  # 5 nearest neighbors

# Create spatial weights
spatial_weights <- nb2listw(knn_neighbors, style = "W")  # Row-standardized weights
```

```{r}
# Moran's I test for EV_Count
moran.test(spatial_data$EV_Count, spatial_weights)

# Moran's I test for residuals of OLS model
ols_model <- lm(EV_Count ~ population + median_income + employed_population + bachelors_or_higher, 
               data = spatial_data)
lm.morantest(ols_model, spatial_weights)
```

```{r}
sar_model <- lagsarlm(EV_Count ~ population + median_income + employed_population + bachelors_or_higher,
                     data = spatial_data, 
                     listw = spatial_weights)
summary(sar_model)
```


```{r}
sem_model <- errorsarlm(EV_Count ~ population + median_income + employed_population + bachelors_or_higher,
                       data = spatial_data, 
                       listw = spatial_weights)
summary(sem_model)
```

```{r}
# Compare AIC values
data.frame(
  Model = c("OLS", "SAR", "SEM"),
  AIC = c(AIC(ols_model), AIC(sar_model), AIC(sem_model))
)

# Likelihood ratio tests
anova(ols_model, sar_model)
anova(ols_model, sem_model)
```

```{r}
library(ggplot2)

# Create a data frame of coefficients
coef_df <- data.frame(
  Variable = names(coef(sar_model))[-1],  # Exclude intercept
  Coefficient = coef(sar_model)[-1],
  Model = "Spatial Lag"
)

# Add SEM coefficients if needed
coef_df <- rbind(coef_df,
                data.frame(
                  Variable = names(coef(sem_model))[-1],
                  Coefficient = coef(sem_model)[-1],
                  Model = "Spatial Error"
                ))

# Plot coefficients
ggplot(coef_df, aes(x = Variable, y = Coefficient, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() +
  labs(title = "Spatial Regression Coefficients",
       subtitle = "Comparison of SAR and SEM Models",
       x = "Predictor Variable",
       y = "Coefficient Estimate") +
  theme_minimal()
```

<<<<<<< Updated upstream
























#######          Q-2           #########
```{r}
ev_population_data <- read.csv("../Dataset/cleaned_ev_data.csv")
summary(ev_population_data)
```


```{r}
ev_population_data <- ev_population_data %>%
  mutate(Make.Model = paste(Make, Model, sep = " "))  # Concatenates with space

# Removing data related to 2025 as we just have data of few months
ev_population_data <- ev_population_data %>%
  filter(Sale.Year != 2025)
```


```{r}
# Install packages if not already installed
required_packages <- c("readr", "dplyr", "lubridate", "ggplot2", "forecast", "prophet")
installed_packages <- rownames(installed.packages())
for (pkg in required_packages) {
  if (!(pkg %in% installed_packages)) {
    install.packages(pkg)
  }
}

# Load libraries
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)
library(prophet)
```



##    Aggregate EV Registrations by Year

```{r}
ev_yearly <- ev_population_data %>%
  group_by(Sale.Year) %>%
  summarise(Total_Registrations = n()) %>%
  arrange(Sale.Year)
```

```{r}
ggplot(ev_yearly, aes(x = factor(Sale.Year), y = Total_Registrations, group = 1)) +
  geom_line(color = "blue") +
  geom_point(color = "darkblue") +
  labs(title = "Total EV Registrations by Year in Washington State",
       x = "Year",
       y = "Number of Registrations") +
  theme_minimal()
```

## Trends in EV Adoption by Make and Model Over Time
```{r}
# Identify top 5 Make+Model combinations
top_models <- ev_population_data %>%
  count(Make.Model, sort = TRUE) %>%
  top_n(5, n) %>%
  pull(Make.Model)

# Filter data for top models
ev_top_models <- ev_population_data %>%
  filter(Make.Model %in% top_models)

# Aggregate registrations by Year and Make_Model
ev_model_yearly <- ev_top_models %>%
  group_by(Sale.Year, Make.Model) %>%
  summarise(Registrations = n()) %>%
  ungroup()

# Plot trends
ggplot(ev_model_yearly, aes(x = factor(Sale.Year), y = Registrations, color = Make.Model)) +
  geom_line() +
  geom_point() +
  labs(title = "EV Registrations Over Time by Make and Model",
       x = "Year",
       y = "Number of Registrations",
       color = "Make and Model") +
  theme_minimal()
```

## Trends in EV Adoption by Vehicle Type Over Time
```{r}
# Aggregate registrations by Year and Electric Vehicle Type
ev_type_yearly <- ev_population_data %>%
  group_by(Sale.Year, Clean.Alternative.Fuel.Vehicle.Type) %>%
  summarise(Registrations = n()) %>%
  ungroup()

# Plot trends
ggplot(ev_type_yearly, aes(x = Sale.Year, y = Registrations, color = Clean.Alternative.Fuel.Vehicle.Type)) +
  geom_line() +
  geom_point() +
  labs(title = "EV Registrations Over Time by Vehicle Type",
       x = "Year",
       y = "Number of Registrations",
       color = "Vehicle Type") +
  theme_minimal()
```

## Compare Growth Rates of Different Make+Model Combinations and Vehicle Types Over Time

```{r}
# Calculate growth rates for top models
ev_model_growth <- ev_model_yearly %>%
  group_by(Make.Model) %>%
  arrange(Sale.Year) %>%
  mutate(Growth_Rate = (Registrations - lag(Registrations)) / lag(Registrations) * 100) %>%
  ungroup()

# Plot growth rates
ggplot(ev_model_growth, aes(x = Sale.Year, y = Growth_Rate, color = Make.Model)) +
  geom_line() +
  geom_point() +
  labs(title = "Year-over-Year Growth Rates by Make and Model",
       x = "Year",
       y = "Growth Rate (%)",
       color = "Make and Model") +
  theme_minimal()
```

## Identify the Most Popular Make+Model Combinations for Each Year
```{r}
# Determine top Make+Model for each year
top_models_per_year <- ev_population_data %>%
  group_by(Sale.Year, Make.Model) %>%
  summarise(Registrations = n()) %>%
  arrange(Sale.Year, desc(Registrations)) %>%
  slice_head(n = 1) %>%
  ungroup()

# Display results
print(top_models_per_year)

summary(ev_data_modified_census_tract)
```



<<<<<<< Updated upstream
## Time-Series Forecasting for Future EV Registrations Using ARIMA
```{r}
# # Create time series object
# ev_ts <- ts(ev_yearly$Total_Registrations, start = min(ev_yearly$Sale.Year), frequency = 1)
# 
# # Fit ARIMA model
# arima_model <- auto.arima(ev_ts)
# 
# # Forecast next 5 years
# arima_forecast <- forecast(arima_model, h = 5)
# 
# # Plot forecast
# autoplot(arima_forecast) +
#   labs(title = "ARIMA Forecast of EV Registrations",
#        x = "Year",
#        y = "Number of Registrations") +
#   theme_minimal()

library(forecast)
library(ggplot2)

# Create time series object
ev_ts <- ts(ev_yearly$Total_Registrations, start = min(ev_yearly$Sale.Year), frequency = 1)

# Fit ARIMA model
arima_model <- auto.arima(ev_ts, seasonal = TRUE, stepwise = FALSE, approximation = FALSE)

# Forecast next 5 years
arima_forecast <- forecast(arima_model, h = 12)

autoplot(arima_forecast)

# # Convert forecast to dataframe for ggplot
# fc_df <- data.frame(
#   Year = seq(from = max(ev_yearly$Sale.Year) + 1, by = 1, length.out = 5),
#   Forecast = arima_forecast$mean,
#   Lo80 = arima_forecast$lower[, 1],
#   Hi80 = arima_forecast$upper[, 1],
#   Lo95 = arima_forecast$lower[, 2],
#   Hi95 = arima_forecast$upper[, 2]
# )
# 
# # Combine with historical data for plotting
# hist_df <- ev_yearly %>%
#   rename(Year = Sale.Year, Registrations = Total_Registrations)
# 
# # Plot
# ggplot() +
#   # Historical data
#   geom_line(data = hist_df, aes(x = Year, y = Registrations), color = "#2c3e50", size = 1.2) +
#   geom_point(data = hist_df, aes(x = Year, y = Registrations), color = "#2980b9", size = 2) +
# 
#   # Forecast line
#   geom_line(data = fc_df, aes(x = Year, y = Forecast), color = "#e74c3c", linetype = "dashed", size = 1.2) +
#   geom_ribbon(data = fc_df, aes(x = Year, ymin = Lo95, ymax = Hi95), fill = "#e74c3c", alpha = 0.15) +
#   geom_ribbon(data = fc_df, aes(x = Year, ymin = Lo80, ymax = Hi80), fill = "#e74c3c", alpha = 0.25) +
#   
#   scale_x_continuous(breaks = seq(min(hist_df$Year), max(fc_df$Year), by = 1)) +
# 
#   # Titles and theme
#   labs(
#     title = "Forecast of EV Registrations (ARIMA Model)",
#     subtitle = "With 80% and 95% Confidence Intervals",
#     x = "Year",
#     y = "Number of Registrations"
#   ) +
#   theme_minimal(base_size = 14) +
#   theme(
#     plot.title = element_text(face = "bold"),
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     legend.position = "none"
#   )
```



```{r}
checkresiduals(arima_model)
```

```{r}
accuracy(arima_model)
```


```{r}
ev_population_data_new <- ev_population_data %>%
  mutate(
    Sale.Date = as.Date(Sale.Date, format = "%B %d %Y"),
    Sale.Month.Year = format(Sale.Date, "%Y-%m")
  )
```


```{r}
library(dplyr)
library(lubridate)
library(ggplot2)

ev_population_data <- ev_population_data %>%
  mutate(
    Sale.Date = as.Date(Sale.Date, format = "%B %d %Y"),
    Sale.Month.Year = format(Sale.Date, "%Y-%m")
  )

# STEP 3: Aggregate registrations per month
monthly_trends <- ev_population_data %>%
  group_by(Sale.Month.Year) %>%
  summarise(EV_Registrations = n()) %>%
  ungroup()

# STEP 4: Plot the trend
ggplot(monthly_trends, aes(x = as.Date(paste0(Sale.Month.Year, "-01")), y = EV_Registrations)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "darkblue") +
  labs(
    title = "Monthly Trend of EV Registrations",
    x = "Month-Year",
    y = "Number of Registrations"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
# STEP 4: Convert to time series (ts object)
# Define start and frequency (monthly = 12)

monthly_trends <- monthly_trends %>%
  mutate(Sale.Month.Year = as.Date(paste0(Sale.Month.Year, "-01")))

# start_year <- year(min(monthly_trends$Sale.Month.Year))
# start_month <- month(min(monthly_trends$Sale.Month.Year))

start_date <- min(monthly_trends$Sale.Month.Year)
start_year <- year(start_date)
start_month <- month(start_date)

ev_ts <- ts(monthly_trends$EV_Registrations,
            start = c(start_year, start_month),
            frequency = 12)

# STEP 5: Fit ARIMA model
arima_model <- auto.arima(ev_ts)

# STEP 6: Forecast next 12 months
forecast_result <- forecast(arima_model, h = 12)

# STEP 7: Plot forecast
autoplot(forecast_result) +
  labs(
    title = "Monthly EV Registration Forecast (ARIMA)",
    x = "Year",
    y = "Number of Registrations"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
summary(arima_model)
```

```{r}
model_order <- arimaorder(arima_model)  # Get p, d, q
model_aic <- AIC(arima_model)

ggplot2::autoplot(forecast(arima_model, h = 12)) +
  labs(
    title = paste0("Monthly EV Registration Forecast - ARIMA(", 
                   model_order[1], ",", model_order[2], ",", model_order[3], ")"),
    subtitle = paste("AIC:", round(model_aic, 2)),
    x = "Year",
    y = "EV Registrations"
  ) +
  theme_minimal(base_size = 14)
```
```



<!-- ## Time-Series Forecasting Using Prophet -->

<!-- ```{r} -->
<!-- # # Prepare data for Prophet -->
<!-- # prophet_data <- ev_yearly %>% -->
<!-- #   rename(ds = Sale.Year, y = Total_Registrations) %>% -->
<!-- #   mutate(ds = as.Date(paste0(ds, "-01-01"))) -->
<!-- #  -->
<!-- # # Fit Prophet model -->
<!-- # prophet_model <- prophet(prophet_data) -->
<!-- #  -->
<!-- # # Create future dataframe for next 5 years -->
<!-- # future <- make_future_dataframe(prophet_model, periods = 5, freq = "year") -->
<!-- #  -->
<!-- # # Forecast -->
<!-- # forecast <- predict(prophet_model, future) -->
<!-- #  -->
<!-- # # Plot forecast -->
<!-- # plot(prophet_model, forecast) + -->
<!-- #   labs(title = "Prophet Forecast of EV Registrations") -->

<!-- library(prophet) -->
<!-- library(ggplot2) -->
<!-- library(dplyr) -->
<!-- library(scales) -->

<!-- # Prepare data for Prophet -->
<!-- prophet_data <- ev_yearly %>% -->
<!--   rename(ds = Sale.Year, y = Total_Registrations) %>% -->
<!--   mutate(ds = as.Date(paste0(ds, "-01-01"))) -->

<!-- # Fit Prophet model -->
<!-- prophet_model <- prophet(prophet_data) -->

<!-- # Create future dataframe for next 5 years -->
<!-- future <- make_future_dataframe(prophet_model, periods = 5, freq = "year") -->

<!-- # Forecast -->
<!-- forecast <- predict(prophet_model, future) -->

<!-- # Merge actual and forecast data for plotting -->
<!-- forecast_plot <- forecast %>% -->
<!--   select(ds, yhat, yhat_lower, yhat_upper) %>% -->
<!--   mutate(ds = as.Date(ds)) %>%     # <-- Ensure it's Date type -->
<!--   left_join(prophet_data, by = "ds") -->

<!-- # Custom ggplot -->
<!-- ggplot(forecast_plot, aes(x = ds)) + -->
<!--   geom_ribbon(aes(ymin = yhat_lower, ymax = yhat_upper), fill = "#cce5ff", alpha = 0.4) + -->
<!--   geom_line(aes(y = yhat), color = "#0073e6", size = 1.2, linetype = "dashed") + -->
<!--   geom_point(aes(y = y), color = "#1c4587", size = 2) + -->
<!--   geom_line(aes(y = y), color = "#1c4587", size = 1) + -->
<!--   scale_x_date(date_breaks = "1 year", date_labels = "%Y") + -->
<!--   labs( -->
<!--     title = "Prophet Forecast of EV Registrations", -->
<!--     x = "Year", -->
<!--     y = "Number of Registrations" -->
<!--   ) + -->
<!--   theme_minimal(base_size = 14) + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 45, hjust = 1), -->
<!--     plot.title = element_text(face = "bold") -->
<!--   ) -->
<!-- ``` -->

=======



>>>>>>> Stashed changes


