
LOADING LIBRARIES
```{r}
library(dplyr)
library(lubridate)
library(stringr)
```

LOADING DATA 
```{r}
ev_data <- read.csv("Electric_Vehicle.csv", stringsAsFactors = FALSE)
```

VIEWING COLUMN NAMES 
```{r}
names(ev_data)
```
DATA CLEANING 

1. Handle missing values
```{r}
ev_data <- ev_data %>%
  mutate(
    Sale.Price = ifelse(Sale.Price == 0, NA, Sale.Price),
    Electric.Range = ifelse(Electric.Range == 0, NA, Electric.Range),
    Odometer.Reading = ifelse(Odometer.Reading == 0, NA, Odometer.Reading)
  ) %>%
  mutate(
    Electric.Range = ifelse(is.na(Electric.Range), median(Electric.Range, na.rm = TRUE), Electric.Range),
    Odometer.Reading = ifelse(is.na(Odometer.Reading), median(Odometer.Reading, na.rm = TRUE), Odometer.Reading)
  )
```

2. Removing Duplicate 
```{r}
ev_data <- ev_data %>%
  distinct(DOL.Vehicle.ID, .keep_all = TRUE)
```

3. Standardize text field 
```{r}
ev_data <- ev_data %>%
  mutate(
    Make = str_to_title(Make),
    Model = str_to_title(Model),
    County = str_to_title(County),
    City = str_to_title(City)
  )
```

4. Convert Categorical Columns to Factors
```{r}
ev_data <- ev_data %>%
  mutate(
    Make = as.factor(Make),
    Model = as.factor(Model),
    City = as.factor(City),
    County = as.factor(County),
    Transaction.Type = as.factor(Transaction.Type)
  )

```

5. Validate Numeric Fields & Outlier Filtering
```{r}
ev_data <- ev_data %>%
  filter(Sale.Price > 1000 & Sale.Price < 200000) %>%
  filter(Model.Year >= 1990 & Model.Year <= year(Sys.Date()))

```

5.Feature Engineering
```{r}
ev_data <- ev_data %>%
  mutate(Transaction.Date = as.Date(Transaction.Date, format = "%B %d %Y"))
ev_data <- ev_data %>%
  mutate(
    Sale.Year = year(Transaction.Date),
    Sale.Month = month(Transaction.Date),
    Sale.Day = day(Transaction.Date),
    Sale.Weekday = weekdays(Transaction.Date)
  )

```

6. Encoding Binary Field 
```{r}
ev_data <- ev_data %>%
  mutate(Electric.Vehicle.Fee.Paid = ifelse(Electric.Vehicle.Fee.Paid == "Yes", 1, 0))
```

7. Split by Transaction Type
```{r}
new_reg <- ev_data %>% filter(Transaction.Type == "New Registration")
title_transfer <- ev_data %>% filter(Transaction.Type == "Title Transfer")
```

8. Dropping extra columns 
```{r}
ev_data <- ev_data %>%
  select(-c(
    Electric.Vehicle.Fee.Paid,
    Transportation.Electrification.Fee.Paid,
    Hybrid.Vehicle.Electrification.Fee.Paid,
    Base.MSRP
  ))

```

SAVING THE CLEANED DATA 
```{r}
write.csv(ev_data, "cleaned_ev_data.csv", row.names = FALSE)
```

DIMENSIONS OF CLEANED DATA
```{r}
dim(ev_data)
```

```{r}
# Count of duplicate VINs
sum(duplicated(ev_data$VIN..1.10.))
sum(duplicated(ev_data$DOL.Vehicle.ID))
```

